---
title: "ROAR-presentation-experiment-3-simulation"
author: "Wanjing Anya Ma"
date:  "`r Sys.Date()`"
output:
  bookdown::html_document2:
    toc: true
    toc_depth: 4
    theme: cosmo
    highlight: tango
editor_options: 
  markdown: 
    wrap: 72
---

# Load Packages

Let's first load the packages that we need for this chapter.

```{r, message=FALSE, warning = FALSE, echo = FALSE}
library(knitr) # for rendering the RMarkdown file
library(tidyverse) # for data wrangling 
library(dplyr)
library(mirt)
library(ggpubr)
library(Metrics)
library(hash)
library(ggplot2)
library(GGally)
library(gridExtra)
library(rstudioapi)
library(viridis)
library(TAM)
library(mgcv)
library(viridis)
library(patchwork)
library(reshape2)
library(catR)
library(janitor)     # for cleaning column names
library(broom.mixed) # for tidying up linear models 
library(emmeans)     # for the joint_tests() function 
library(lme4)        # for linear mixed effects models
library(performance) # for assessing model performance
library(see)         # for assessing model performance
```

## Settings

```{r echo = FALSE}
# sets how code looks in knitted document
opts_chunk$set(comment = "")

# suppresses warning about grouping 
options(dplyr.summarise.inform = F)
```

```{r}
timeConvert <- c(
  "80" ="T1",
  "160"= "T2",
  "350"="T3",
  "700"= "T4",
  "infinite"= "T5"
)
```

## load corpus

```{r}
df.new.item.bank <- read.csv("corpus/item_bank_with_slow_items_v6.csv")
```

# Sim: Benefit of two banks

what is the theta and st error to transition? what is the theta and st to stop early?
```{r}
# inverse logit
inv_logit <- function(x) {
  0.5 + 0.5 * (exp(x) / (1 + exp(x)))
}

# simulate responses
func.create.response <- function(b, th, np, ni) {
  th.mat <- matrix(th, np, ni, byrow = FALSE)
  b.mat <- matrix(rep(b, np), nrow = np, byrow = TRUE)
  pr <- inv_logit(th.mat - b.mat)
  resp <- pr
  for (i in 1:ncol(resp)) {
    resp[, i] <- rbinom(nrow(resp), 1, resp[, i])
  }
  return(as.data.frame(resp))
}

```

```{r}
func.twoPhasePreSim <- function(item.bank, thetas.true,
                                easyCorpus = "en-corpusNew-easy",
                                nEasyStart = 0, maxItems = 25) {
  
  # split easy vs hard
  easyBank_full <- item.bank %>% filter(corpusId == easyCorpus)
  
  easyBank <- as.matrix(easyBank_full %>% select(a,b,c,d))
  
  np <- length(thetas.true)
  
  # simulate full responses (phase 1 and 2 separately)
  resp_easy_full <- func.create.response(easyBank_full$b, thetas.true, np, nrow(easyBank_full))
  
  test_ctrl  <- list(method = "ML", itemSelect = "MFI", infoType = "Fisher", range = c(-6,6))
  final_ctrl <- list(method = "ML")
  
  results_all <- list()
  
  for (i in seq_len(np)) {
    
    
    pid <- i
    th_true <- thetas.true[i]
    
    ## Phase 1: Easy items
    start_easy <- list(fixItems = sample(seq_len(nrow(easyBank)), 1), nrItems = 1, theta = 0)
    
    easyBank <- balanced_item_bank %>% filter(corpusId == "en-corpusNew-easy")
    cbGroup_easy <- as.character(easyBank$group)
    cbList <- list(names = c("real", "pseudo"), props = c(0.5, 0.5))
    
    stop_ctl <- list(rule = 'length',thr = maxItems)
    final <- list(method = 'ML')
    
    res_easy <- randomCAT(
      itemBank  = as.matrix(easyBank[, c("a","b","c","d", "group")]),
      responses = as.numeric(resp_easy_full[i, ]),
      start     = start_easy,
      test      = test_ctrl,
      stop = stop_ctl,
      final     = final_ctrl,
      cbControl = cbList
      )
    
    pCorrect_easy <- mean(res_easy$pattern)
    df_easy <- data.frame(
      pid = pid,
      trialNumTotal = seq(1,maxItems,1),
      item = easyBank_full$word[res_easy$testItems],
      a = easyBank_full$a[res_easy$testItems],
      b = easyBank_full$b[res_easy$testItems],
      c = easyBank_full$c[res_easy$testItems],
      d = easyBank_full$d[res_easy$testItems],
      resp = res_easy$pattern,
      corpusId = easyBank_full$corpusId[res_easy$testItems],
      realpseudo = easyBank_full$group[res_easy$testItems],
      thetaEstimate = res_easy$thetaProv,
      thetaSE = res_easy$seProv,
      thetaTrue = th_true,
      phase = "easy"
    )
    
    results_all[[i]] <- rbind(df_easy)
    
  }
  
  return(list(
    results_all = bind_rows(results_all),
    resp_easy_full = resp_easy_full
  ))
}
```

```{r}
func.twoPhasePreSim.hard <- function(item.bank, thetas.true,
                                easyCorpus = "validated",
                                nEasyStart = 0, maxItems = 25) {
  
  # split easy vs hard
  easyBank_full <- item.bank %>% filter(corpusId == easyCorpus)
  
  easyBank <- as.matrix(easyBank_full %>% select(a,b,c,d))
  
  np <- length(thetas.true)
  
  # simulate full responses (phase 1 and 2 separately)
  resp_easy_full <- func.create.response(easyBank_full$b, thetas.true, np, nrow(easyBank_full))
  
  
  
  test_ctrl  <- list(method = "ML", itemSelect = "MFI", infoType = "Fisher", range = c(-6,6))
  final_ctrl <- list(method = "ML")
  
  results_all <- list()
  
  for (i in seq_len(np)) {
    if (i %% 100 == 0) {
      print(i)  # will print 30
    }
    
    
    pid <- i
    th_true <- thetas.true[i]
    
    start_easy <- list(fixItems = sample(seq_len(nrow(easyBank)), 1), nrItems = 1, theta = 0)
  
    easyBank <- balanced_item_bank %>% filter(corpusId != "en-corpusNew-easy")
    cbGroup_easy <- as.character(easyBank$group)
    cbList <- list(names = c("real", "pseudo"), props = c(0.5, 0.5))
    
    stop_ctl <- list(rule = 'length',thr = maxItems)
    
    
    final <- list(method = 'ML')
    
    res_easy <- randomCAT(
      itemBank  = as.matrix(easyBank[, c("a","b","c","d", "group")]),
      responses = as.numeric(resp_easy_full[i, ]),
      start     = start_easy,
      test      = test_ctrl,
      stop      = stop_ctl,
      final     = final_ctrl,
      cbControl = cbList
      )
    
    pCorrect_easy <- mean(res_easy$pattern)
    df_easy <- data.frame(
      pid = pid,
      trialNumTotal = seq(1,maxItems,1),
      item = easyBank_full$word[res_easy$testItems],
      a = easyBank_full$a[res_easy$testItems],
      b = easyBank_full$b[res_easy$testItems],
      c = easyBank_full$c[res_easy$testItems],
      d = easyBank_full$d[res_easy$testItems],
      resp = res_easy$pattern,
      corpusId = easyBank_full$corpusId[res_easy$testItems],
      realpseudo = easyBank_full$group[res_easy$testItems],
      thetaEstimate = res_easy$thetaProv,
      thetaSE = res_easy$seProv,
      thetaTrue = th_true,
      phase = "hard"
    )
    
    results_all[[i]] <- rbind(df_easy)
  }
  
  bind_rows(results_all)
}
```

```{r}
balanced_item_bank <- df.new.item.bank %>%
  group_by(corpusId) %>%
  group_modify(~ {
    # .x is the data for this corpusId
    min_n <- min(table(.x$realpseudo))
    
    # sample min_n items from each realpseudo group
    .x %>%
      group_by(realpseudo) %>%
      slice_sample(n = min_n)
  }) %>%
  ungroup() %>% 
  rename(group = realpseudo) %>% 
  arrange()
```

```{r}
theta_vals <- seq(-6, 6, by = 0.1)
theta_true <- rep(theta_vals, each = 10)  # 10 participants per theta

sim_results_bundle <- func.twoPhasePreSim(
  item.bank = balanced_item_bank,
  thetas.true = theta_true,
  easyCorpus = "en-corpusNew-easy",
  nEasyStart = 0,
  maxItems = 84
)
```

```{r}
sim_results <- sim_results_bundle$results_all

sim_results_resp <- sim_results_bundle$resp_easy_full
```

```{r}
sim_results.hard <- func.twoPhasePreSim.hard(
  item.bank = balanced_item_bank,
  thetas.true = theta_true,
  easyCorpus = "validated",
  nEasyStart = 0,
  maxItems = 84
)
```


```{r}
df_eval <- sim_results %>%
  rbind(sim_results.hard) %>% 
  mutate(theta_group = round(thetaTrue)) %>% 
  group_by(trialNumTotal, theta_group, phase) %>%
  dplyr::summarise(n = n(), sem = mean(thetaSE), 
                     reliability = empirical_rxx(as.matrix(tibble(F1 = thetaEstimate, SE_F1 = thetaSE))),
                     mse = Metrics :: rmse(thetaTrue, thetaEstimate),
                     bias = Metrics :: bias(thetaTrue, thetaEstimate))
```

```{r}
df_eval_v2 <- sim_results %>%
  rbind(sim_results.hard) %>% 
  mutate(theta_group = round(thetaTrue)) %>% 
  group_by(trialNumTotal, theta_group, phase, thetaTrue) %>%
  dplyr::summarise(n = n(), sem = mean(thetaSE), 
                     reliability = empirical_rxx(as.matrix(tibble(F1 = thetaEstimate, SE_F1 = thetaSE))),
                     mse = Metrics :: rmse(thetaTrue, thetaEstimate),
                     bias = Metrics :: bias(thetaTrue, thetaEstimate))

ggplot(df_eval_v2 %>% 
  filter(trialNumTotal == 84), aes(x = thetaTrue, y = sem, color = phase)) +
  geom_smooth(size = 0.5) +
  theme_minimal() +
  labs(title = "SEM by Theta Estimate",
       x = "Theta Estimate", y = "SEM") + ylim(0, 5) 
```

```{r}
# Plot MSE
p_mse <- ggplot(df_eval, aes(x = trialNumTotal, y = mse, color = phase)) +
  geom_point(size = 0.5) +
  theme_minimal() +
  labs(title = "MSE by Theta Estimate",
       x = "Theta Estimate", y = "Mean Squared Error") + ylim(0, 5) + facet_wrap(~theta_group)

# Plot Bias
p_bias <- ggplot(df_eval, aes(x = trialNumTotal, y = bias, color = phase)) +
  geom_point() +
  theme_minimal() +
  labs(title = "Bias by Theta Estimate",
       x = "Theta Estimate", y = "Bias") + facet_wrap(~theta_group)


# df_eval_all <- sim_results %>%
#   rbind(sim_results.hard) %>% 
#   mutate(theta_group = round(thetaTrue)) %>% 
#   group_by(trialNumTotal, phase) %>%
#   dplyr::summarise(n = n(), sem = mean(thetaSE), 
#                      reliability = empirical_rxx(as.matrix(tibble(F1 = thetaEstimate, SE_F1 = thetaSE))),
#                      mse = Metrics :: rmse(thetaTrue, thetaEstimate), 
#                      cor = cor(thetaTrue, thetaEstimate), 
#                      bias = Metrics :: bias(thetaTrue, thetaEstimate))

# Plot Reliability
p_rel <- ggplot(df_eval, aes(x = trialNumTotal, y = reliability, color = phase)) +
  geom_point(size = 0.5) +
  theme_minimal() +
  labs(title = "Reliability by Theta Estimate",
       x = "Theta Estimate", y = "Reliability")+ ylim(0, 1)+ facet_wrap(~theta_group)

p_sem <- ggplot(df_eval, aes(x = trialNumTotal, y = sem, color = phase)) +
  geom_point(size = 0.5) +
  theme_minimal() +
  labs(title = "SEM by Theta Estimate",
       x = "Theta Estimate", y = "sem")+ ylim(0, 5) + facet_wrap(~theta_group)

# Show plots
p_mse
p_bias
p_rel
p_sem
```

# Sim: Transition Rule
run participants (theta -6, 6) 

## How to transition?

```{r}
easy_items <- balanced_item_bank %>% 
  filter(corpusId == "en-corpusNew-easy") 

sim_results_resp_name <- sim_results_resp %>%
  dplyr::rename_with(~ easy_items$word, .cols = everything())

df.sim.base <- sim_results_resp_name %>% 
  add_column(pid = seq(1, 1210, 1)) %>% 
  relocate(pid) %>% 
  add_column(theta_true = theta_true) %>% 
  tidyr::pivot_longer(
    cols = -c(pid, theta_true),
    names_to = "item",
    values_to = "resp"
  ) %>% 
  right_join(sim_results) 
```

```{r}
find_first_qualified_trial <- function(df,
                                       n_in_row = 5,
                                       theta_threshold = -3,
                                       resp_col = "resp",
                                       theta_col = "thetaEstimate",
                                       se_col = "thetaSE",
                                       trial_col = "trialNumTotal",
                                       pid_col = "pid",
                                       theta_true_col = "theta_true") {
  
  resp_sym  <- rlang::sym(resp_col)
  theta_sym <- rlang::sym(theta_col)
  se_sym    <- rlang::sym(se_col)
  trial_sym <- rlang::sym(trial_col)
  pid_sym   <- rlang::sym(pid_col)
  true_sym  <- rlang::sym(theta_true_col)
  
  df %>%
    arrange(!!pid_sym, !!trial_sym) %>%
    group_by(!!pid_sym) %>%
    mutate(
      # rolling sum of last n_in_row responses
      roll_sum = Reduce(
        `+`,
        lapply(0:(n_in_row - 1), function(k)
          lag(!!resp_sym, k, default = 0))
      ),
      
      hit_now = roll_sum == n_in_row,
      ever_hit = cumany(hit_now),
      
      qualified = ever_hit & (!!theta_sym >= theta_threshold)
    ) %>%
    summarize(
      !!trial_col := if(any(qualified)) {
        first(which(qualified))
      } else {
        n()
      },
      
      !!theta_col := (!!theta_sym)[if(any(qualified)) first(which(qualified)) else n()],
      !!se_col    := (!!se_sym)[if(any(qualified)) first(which(qualified)) else n()],
      !!theta_true_col := first(!!true_sym),
      
      .groups = "drop"
    )
}
```

```{r}
df.first4.2 <- find_first_qualified_trial(
  df.sim.base,
  n_in_row = 4,
  theta_threshold = -2
)

df.first5.2 <- find_first_qualified_trial(
  df.sim.base,
  n_in_row = 5,
  theta_threshold = -2
)

df.first4 <- find_first_qualified_trial(
  df.sim.base,
  n_in_row = 4,
  theta_threshold = -3
)

df.first5 <- find_first_qualified_trial(
  df.sim.base,
  n_in_row = 5,
  theta_threshold = -3
)
```


```{r}
ggplot(df.first4.2, aes(x = theta_true, y = trialNumTotal, color = thetaEstimate)) +
  geom_point(alpha = 0.6) +
  geom_smooth(method = "loess", se = TRUE) +
  labs(
    x = "True Theta",
    y = "Trial Number of First 3-in-a-Row & Theta > -3",
    title = "Transitioning Pattern"
  ) +
  theme_minimal()

ggplot(df.first5.2, aes(x = theta_true, y = trialNumTotal, color = thetaEstimate)) +
  geom_point(alpha = 0.6) +
  geom_smooth(method = "loess", se = TRUE) +
  labs(
    x = "True Theta",
    y = "Trial Number of First 3-in-a-Row & Theta > -3",
    title = "Transitioning Pattern"
  ) +
  theme_minimal()

ggplot(df.first4, aes(x = theta_true, y = trialNumTotal, color = thetaEstimate)) +
  geom_point(alpha = 0.6) +
  geom_smooth(method = "loess", se = TRUE) +
  labs(
    x = "True Theta",
    y = "Trial Number of First 4-in-a-Row & Theta > -3",
    title = "Transitioning Pattern"
  ) +
  theme_minimal()

ggplot(df.first5, aes(x = theta_true, y = trialNumTotal, color = thetaEstimate)) +
  geom_point(alpha = 0.6) +
  geom_smooth(method = "loess", se = TRUE) +
  labs(
    x = "True Theta",
    y = "Trial Number of First 5-in-a-Row & Theta > -3",
    title = "Transitioning Pattern"
  ) +
  theme_minimal()
```

- create SEM curve that is the widest. 
- options: 3 in a row, 4 in a row,  5 in a row. 

stopping rule: 
- true label: chance performer vs. non-chance performer 
- how many trials do you need? 

## classification eval
```{r}

make_classification_metrics <- function(df,
                                        theta_cut = -3,
                                        trial_cut = 20) {
  safe_div <- function(num, den) ifelse(den == 0, NA, num / den)
  
  df2 <- df %>%
    dplyr::mutate(
      true_class = ifelse(theta_true < theta_cut, 0, 1),
      pred_class = ifelse(trialNumTotal < trial_cut, 1, 0)
    )
  
  cm <- table(True = df2$true_class,
              Pred = df2$pred_class)
  
  
  # Extract cells
  TP <- cm["1","1"]
  TN <- cm["0","0"]
  FP <- cm["0","1"]
  FN <- cm["1","0"]
  
  # Metrics
  accuracy  <- (TP + TN) / sum(cm)
  precision <- TP / (TP + FP)
  recall    <- TP / (TP + FN)
  
  f1 <- safe_div(2 * precision * recall,
                 precision + recall)
  
  list(
    confusion_matrix = cm,
    accuracy  = accuracy,
    precision = precision,
    recall    = recall,
    f1 = f1
  )
}
```


```{r}
# Run metrics for each dataset
metrics_first5 <- make_classification_metrics(df.first5)
metrics_first4 <- make_classification_metrics(df.first4)
metrics_first4_2 <- make_classification_metrics(df.first4.2)
metrics_first5_2 <- make_classification_metrics(df.first5.2)

# Convert to tibble
report_table <- bind_rows(
  tibble(
    dataset = "a: first 4 in a row with -3",
    accuracy = metrics_first4$accuracy,
    precision = metrics_first4$precision,
    recall = metrics_first4$recall,
    balanced_f1 = metrics_first4$f1
  ),
  tibble(
    dataset = "b: first 4 in a row with -2",
    accuracy = metrics_first4_2$accuracy,
    precision = metrics_first4_2$precision,
    recall = metrics_first4_2$recall,
    balanced_f1 = metrics_first4_2$f1
  ),
  tibble(
    dataset = "c: first 5 in a row with -3",
    accuracy = metrics_first5$accuracy,
    precision = metrics_first5$precision,
    recall = metrics_first5$recall,
    balanced_f1 = metrics_first5$f1
  ),
  tibble(
    dataset = "d: first 5 in a row with -2",
    accuracy = metrics_first5_2$accuracy,
    precision = metrics_first5_2$precision,
    recall = metrics_first5_2$recall,
    balanced_f1 = metrics_first5_2$f1
  ),
)

report_table
write.csv(report_table, "results/study2_transition_rule_simulation_classification_results.csv")
```
limitation of the simulation. 

## SEM eval

```{r}
func.twoPhasePreSim.later <- function(
  item.bank,
  thetas.true,
  df,
  easyCorpus = "validated",
  nEasyStart = 0,
  maxItems = 84
) {
  
  ## ---- split easy vs hard ----
  easyBank_full <- item.bank %>%
    filter(corpusId == easyCorpus)
  
  easyBank <- as.matrix(
    easyBank_full %>% select(a, b, c, d)
  )
  
  np <- length(thetas.true)
  
  ## ---- simulate responses ----
  resp_easy_full <- func.create.response(
    easyBank_full$b,
    thetas.true,
    np,
    nrow(easyBank_full)
  )
  
  test_ctrl  <- list(
    method = "ML",
    itemSelect = "MFI",
    infoType = "Fisher",
    range = c(-6, 6)
  )
  
  final_ctrl <- list(method = "ML")
  
  results_all <- vector("list", np)
  
  ## =====================================================
  ##                    MAIN LOOP
  ## =====================================================
  for (i in seq_len(np)) {
    
    pid <- i
    th_true <- thetas.true[i]
    
    ## ---- pull prior state from df (or initialize) ----
    df_pid <- df %>% filter(pid == !!pid)
    
    if (nrow(df_pid) == 0) {
      theta_start <- 0
      trialNumTotal <- 0
    } else {
      theta_start   <- df_pid$thetaEstimate[1]
      trialNumTotal <- df_pid$trialNumTotal[1]
    }
    
    if (trialNumTotal == 84){
      next
    }
    
    ## ---- starting state ----
    start_easy <- list(
      fixItems = sample(seq_len(nrow(easyBank)), 1),
      nrItems  = 1,
      theta    = theta_start
    )
    
    ## ---- content balancing ----
    easyBank_cb <- balanced_item_bank %>%
      filter(corpusId != "en-corpusNew-easy")
    
    cbList <- list(
      names = c("real", "pseudo"),
      props = c(0.5, 0.5)
    )
    
    ## ---- stop rule respects prior items ----
    stop_ctl <- list(
      rule = "length",
      thr  = maxItems - trialNumTotal
    )
    
    ## ---- run CAT ----
    res_easy <- randomCAT(
      itemBank  = as.matrix(
        easyBank_cb[, c("a", "b", "c", "d", "group")]
      ),
      responses = as.numeric(resp_easy_full[i, ]),
      start     = start_easy,
      test      = test_ctrl,
      stop      = stop_ctl,
      final     = final_ctrl,
      cbControl = cbList
    )
    
    pCorrect_easy <- mean(res_easy$pattern)
    df_easy <- data.frame(
      pid = pid,
      trialNumTotal = seq(trialNumTotal+1,maxItems,1),
      item = easyBank_full$word[res_easy$testItems],
      a = easyBank_full$a[res_easy$testItems],
      b = easyBank_full$b[res_easy$testItems],
      c = easyBank_full$c[res_easy$testItems],
      d = easyBank_full$d[res_easy$testItems],
      resp = res_easy$pattern,
      corpusId = easyBank_full$corpusId[res_easy$testItems],
      realpseudo = easyBank_full$group[res_easy$testItems],
      thetaEstimate = res_easy$thetaProv,
      thetaSE = res_easy$seProv,
      thetaTrue = th_true,
      phase = "hard"
    )
    
    results_all[[i]] <- rbind(df_easy)
    
  }
  
  ## ---- return updated df ----
  bind_rows(results_all)
}
```

```{r}

sim_results.later.5 <- func.twoPhasePreSim.later(item.bank = balanced_item_bank,
  thetas.true = theta_true,
  df.first5.2,
  easyCorpus = "validated",
  nEasyStart = 0,
  maxItems = 84)
```

```{r}
sim_results.later.4 <- func.twoPhasePreSim.later(item.bank = balanced_item_bank,
  thetas.true = theta_true,
  df.first4.2,
  easyCorpus = "validated",
  nEasyStart = 0,
  maxItems = 84)
```


```{r}
# hi.new.sim.select <- sim_results.later %>% 
#   select(pid, thetaTrue, item, resp, trialNumTotal, thetaEstimate, thetaSE, phase) %>% 
#   rbind(df.sim.base %>% select(pid, thetaTrue, item, resp, trialNumTotal, thetaEstimate, thetaSE, phase)) %>% 
#   group_by(pid, trialNumTotal, thetaTrue) %>% 
#   summarise(phase = max(phase)) %>% 
#   filter(trialNumTotal == 84) %>% 
#   mutate(keep = 1)
# 
# df_eval_new <- sim_results.later %>% 
#   select(pid, thetaTrue, item, resp, trialNumTotal, thetaEstimate, thetaSE, phase) %>% 
#   rbind(df.sim.base %>% select(pid, thetaTrue, item, resp, trialNumTotal, thetaEstimate, thetaSE, phase)) %>% 
#   filter(trialNumTotal == 84) %>% 
#   left_join(hi.new.sim.select) %>% 
#   filter(keep == 1) %>% 
#   rename(sem = thetaSE) %>% 
#   mutate(phase = "new")
# 
# df_compare <- sim_results %>%
#   rbind(sim_results.hard) %>% 
#   mutate(theta_group = round(thetaTrue)) %>% 
#   filter(trialNumTotal == 84) %>% 
#   select(thetaTrue, phase, thetaSE, thetaEstimate) %>% 
#   rename(sem = thetaSE) %>% 
#   rbind(df_eval_new %>% select(thetaTrue, sem, phase, thetaEstimate))

```


```{r}
df.new.simulated.rule.5 <- sim_results.later.5 %>% 
  select(pid, thetaTrue, item, resp, trialNumTotal, thetaEstimate, thetaSE, phase) %>% 
  rbind(df.sim.base %>% select(pid, thetaTrue, item, resp, trialNumTotal, thetaEstimate, thetaSE, phase)) %>% 
  group_by(pid, trialNumTotal, thetaTrue) %>% 
  summarise(phase = max(phase)) %>% 
  mutate(keep = 1) %>% right_join(sim_results.later.5 %>% 
  select(pid, thetaTrue, item, resp, trialNumTotal, thetaEstimate, thetaSE, phase) %>% 
  rbind(df.sim.base %>% select(pid, thetaTrue, item, resp, trialNumTotal, thetaEstimate, thetaSE, phase))) %>%
  filter(keep == 1)

df.new.simulated.rule.4 <- sim_results.later.4 %>% 
  select(pid, thetaTrue, item, resp, trialNumTotal, thetaEstimate, thetaSE, phase) %>% 
  rbind(df.sim.base %>% select(pid, thetaTrue, item, resp, trialNumTotal, thetaEstimate, thetaSE, phase)) %>% 
  group_by(pid, trialNumTotal, thetaTrue) %>% 
  summarise(phase = max(phase)) %>% 
  mutate(keep = 1) %>% right_join(sim_results.later.4 %>% 
  select(pid, thetaTrue, item, resp, trialNumTotal, thetaEstimate, thetaSE, phase) %>% 
  rbind(df.sim.base %>% select(pid, thetaTrue, item, resp, trialNumTotal, thetaEstimate, thetaSE, phase))) %>%
  filter(keep == 1)
```
```{r}
df.new.simulated.rule %>% 
  left_join(df.new.item.bank %>% select(word, a, b, c, d) %>% rename(item = word))
```
```{r}
build_scoring_model_3pl <- function(item_bank){
  stopifnot(all(c("a","b","c") %in% names(item_bank)))
  n_items <- nrow(item_bank)

  # dummy numeric data (not used for fitting)
  dat <- matrix(
  sample(c(0, 1), size = 100 * n_items, replace = TRUE),
  nrow = 100,
  ncol = n_items
)
  

colnames(dat) <- paste0("item_", seq_len(n_items))

  # create a mirt parameter template, then fill it
  mod0 <- mirt(dat, 1, itemtype = "3PL", verbose = FALSE)
  pars <- mod2values(mod0)
  print(1)

  # fill item parameters (mirt uses a1 and d; d = -a*b)
  pars$value[pars$name == "a1"] <- item_bank$a
  pars$value[pars$name == "d"]  <- -item_bank$a * item_bank$b
  pars$value[pars$name == "g"]  <- item_bank$c
  pars$value[pars$name == "u"]  <- 1
  print(2)

  # lock them (so nothing is estimated)
  pars$est <- FALSE

  # build the scoring model from fixed parameters
  mirt(dat, 1, itemtype = "3PL", pars = pars, pars.only = TRUE, verbose = FALSE)
}

score_with_item_bank <- function(resp_wide, item_bank, method="ML", quadpts=61){

  pars_mirt <- data.frame(
    a1 = item_bank$a,
    d  = -item_bank$a * item_bank$b,
    g  = item_bank$c,
    u  = 1
  )

  n_items <- nrow(item_bank)
  dummy_resp <- matrix(1, nrow = 1, ncol = n_items)
  colnames(dummy_resp) <- paste0("Item", seq_len(n_items))
  
  
  scoring_model <- build_scoring_model_3pl(item_bank)
  print(3)

resp_matrix <- resp_wide[, -1, drop=FALSE]
resp_matrix[] <- lapply(resp_matrix, function(x) as.numeric(as.character(x)))
resp_matrix <- as.matrix(resp_matrix)
colnames(resp_matrix) <- colnames(extract.mirt(scoring_model, "data"))

fs <- fscores(scoring_model, response.pattern = resp_matrix,
              method = method, full.scores.SE = TRUE, quadpts = quadpts, max_theta = 6, min_theta = -6)


  data.frame(pid = resp_wide[[1]], theta = fs[, "F1"], theta_se = fs[, "SE_F1"])
}
```

```{r}
df.new.item.bank.orderd <- df.new.item.bank %>% 
  arrange(word)
```


```{r}
df_resp_new.4 <- df.new.simulated.rule.4 %>% 
  left_join(df.new.item.bank %>% select(word, a, b, c, d) %>% rename(item = word)) %>% 
  select(pid, resp, item)

df_resp_new.5 <- df.new.simulated.rule.5 %>% 
  left_join(df.new.item.bank %>% select(word, a, b, c, d) %>% rename(item = word)) %>% 
  select(pid, resp, item)

```

```{r}
make_resp_wide <- function(df, item_bank, fake_pid = -1) {
  
  fake_df <- data.frame(
    pid  = fake_pid,
    resp = 1,
    item = item_bank$word
  )
  
  df_full <- dplyr::bind_rows(df, fake_df)
  
  resp_wide <- df_full %>%
    dplyr::ungroup() %>%
    dplyr::select(pid, item, resp) %>%
    tidyr::pivot_wider(
      names_from  = item,
      values_from = resp,
      values_fill = NA
    ) %>%
    dplyr::arrange(pid) %>%
    dplyr::filter(pid != fake_pid)
  
  # ensure exact item order
  resp_wide <- resp_wide[, c("pid", item_bank$word)]
  
  return(resp_wide)
}

resp_wide_hard <- make_resp_wide(sim_results.hard, df.new.item.bank.orderd)

resp_wide_easy <- make_resp_wide(sim_results, df.new.item.bank.orderd)

```

```{r}
resp_wide_new_4 <- make_resp_wide(df_resp_new.4 , df.new.item.bank.orderd)

resp_wide_new_5 <- make_resp_wide(df_resp_new.5, df.new.item.bank.orderd)
```
```{r}
theta_df_new_4 <- score_with_item_bank(resp_wide_new_4,
                                 df.new.item.bank.orderd,
                                 method = "ML")

theta_df_new_5 <- score_with_item_bank(resp_wide_new_5,
                                 df.new.item.bank.orderd,
                                 method = "ML")
```

```{r}
theta_df_hard <- score_with_item_bank(resp_wide_hard,
                                 df.new.item.bank.orderd,
                                 method = "ML") 

theta_df_easy <- score_with_item_bank(resp_wide_easy,
                                 df.new.item.bank.orderd,
                                 method = "ML") 
```

```{r}
df_eval_new_4 <- hi.new.sim.select %>% 
  select(pid, thetaTrue) %>% 
  add_column(thetaEstimate = theta_df_new_4$theta, sem = theta_df_new_4$theta_se) %>% 
  mutate(phase = "Adaptive timing") %>% 
  ungroup()

df_eval_new_5 <- hi.new.sim.select %>% 
  select(pid, thetaTrue) %>% 
  add_column(thetaEstimate = theta_df_new_5$theta, sem = theta_df_new_5$theta_se) %>% 
  mutate(phase = "Adaptive timing") %>% 
  ungroup()

df_eval_hard <- hi.new.sim.select %>% 
  select(pid, thetaTrue) %>% 
  add_column(thetaEstimate = theta_df_hard$theta, sem = theta_df_hard$theta_se) %>% 
  mutate(phase = "350 ms") %>% 
  ungroup()

df_eval_easy <- hi.new.sim.select %>% 
  select(pid, thetaTrue) %>% 
  add_column(thetaEstimate = theta_df_easy$theta, sem = theta_df_easy$theta_se) %>% 
  mutate(phase = "Infinite") %>% 
  ungroup()


df_compare <- df_eval_new_4 %>% 
  rbind(df_eval_hard) %>% 
  rbind(df_eval_easy)

df.plot <- df_compare %>% 
  filter(!is.na(sem)) %>% 
  group_by(thetaTrue, phase) %>% 
  summarise(sem = mean(sem, na.rm = TRUE), mse = Metrics :: mse(thetaTrue, thetaEstimate)) %>% 
  rename(Test = phase) %>% 
  mutate(
    Test = factor(Test, levels = c("350 ms", "Infinite", "Adaptive timing"))
  )
  

```

```{r}
g.study3.sem <- ggplot(df.plot, aes(x = thetaTrue, y = sem, color = Test, fill = Test, group= Test)) +
  geom_smooth(size = 1) +
  scale_color_manual(values = c( "black","#E05A1D", "#8F993E")) +
  scale_fill_manual(values = c("black", "#E05A1D", "#8F993E")) +
  labs(title = "Standard error of measurement",
       x = "Theta estimate", y = "Standard error of measurement") + ylim(0, 3) + theme(legend.position = "None") +theme(legend.text = element_text(size = 12), 
        legend.title = element_text(size = 14),
        title = element_text(size = 14),
        axis.text = element_text(size = 12),
        axis.title = element_text(size = 14, margin = margin(t = 10)))

g.study3.mse <-ggplot(df.plot, aes(x = thetaTrue, y = mse, color = Test, fill = Test, group= Test)) +
  geom_smooth(size = 1) +
  scale_color_manual(values = c("black", "#E05A1D", "#8F993E")) +
  scale_fill_manual(values = c("black", "#E05A1D", "#8F993E")) +
  labs(title = "Mean squared error",
       x = "Theta estimate", y = "Mean squared error") + ylim(0, 2) + theme(legend.position = "none") + theme(legend.text = element_text(size = 12),                                                                                                    title = element_text(size = 14),
        legend.title = element_text(size = 14),
        axis.text = element_text(size = 12),
        axis.title = element_text(size = 14, margin = margin(t = 10)))


(g.study3.sem + g.study3.mse) +
  plot_layout(guides = "collect") &
  theme(legend.position = "bottom")

ggsave("plots/roar_presentation_study3_transition_rule.pdf", width = 8, height = 4, units = "in")
```
```{r}
g.study3.sem.demo <- ggplot(df.plot %>% filter(Test!= "Adaptive timing"), aes(x = thetaTrue, y = sem, color = Test, fill = Test, group= Test)) +
  geom_smooth(size = 1) +
  scale_color_manual(values = c( "black","#E05A1D", "#8F993E")) +
  scale_fill_manual(values = c("black", "#E05A1D", "#8F993E")) +
  labs(title = "Standard error of measurement",
       x = "Theta estimate", y = "Standard error of measurement") + ylim(0, 3) + theme(legend.position = "None") +theme(legend.text = element_text(size = 12), 
        legend.title = element_text(size = 14),
        title = element_text(size = 14),
        axis.text = element_text(size = 12),
        axis.title = element_text(size = 14, margin = margin(t = 10)))

g.study3.mse.demo <-ggplot(df.plot %>% filter(Test!= "Adaptive timing"), aes(x = thetaTrue, y = mse, color = Test, fill = Test, group= Test)) +
  geom_smooth(size = 1) +
  scale_color_manual(values = c("black", "#E05A1D", "#8F993E")) +
  scale_fill_manual(values = c("black", "#E05A1D", "#8F993E")) +
  labs(title = "Mean squared error",
       x = "Theta estimate", y = "Mean squared error") + ylim(0, 2) + theme(legend.position = "none") + theme(legend.text = element_text(size = 12),                                                                                                    title = element_text(size = 14),
        legend.title = element_text(size = 14),
        axis.text = element_text(size = 12),
        axis.title = element_text(size = 14, margin = margin(t = 10)))


(g.study3.sem.demo + g.study3.mse.demo) +
  plot_layout(guides = "collect") &
  theme(legend.position = "bottom")

#ggsave("plots/roar_presentation_study3_transition_rule.pdf", width = 8, height = 4, units = "in")
```

# Sim: Early Stopping

## Theoretical Simulations 
Idea: simulate a mix of students who are (a) at floor or careless respondents, or (b) students can read and not careless. 

```{r}
theta_true <- sample(df.thetas$trueTheta, size = 5000, replace = TRUE)
```

```{r}
sum(theta_true < -5)
```

```{r}
sim_results_bundle <- func.twoPhasePreSim(
  item.bank = balanced_item_bank,
  thetas.true = theta_true,
  easyCorpus = "en-corpusNew-easy",
  nEasyStart = 0,
  maxItems = 84
)
```


```{r}
hi <- sim_results_bundle$results_all
```

```{r}
hi
```

```{r}

ggplot(hi %>% 
  filter(thetaTrue == -6, trialNumTotal == 84),
       aes(x = thetaEstimate,
           y = thetaSE)) +
  geom_point(se = FALSE, alpha = 0.2)


ggplot(hi %>% 
  filter(thetaTrue == -6, trialNumTotal == 84),
       aes(x = thetaEstimate)) +
  geom_histogram(se = FALSE, alpha = 0.5, bins = 20)
```
```{r}

```

```{r}
hi %>% 
  filter(trialNumTotal <= 40, trialNumTotal > 20) %>% 
  group_by(pid, thetaTrue) %>% 
  summarise(acc = mean(resp),max_theta =  max(thetaEstimate)) %>% 
  mutate(early_stop = ifelse(max_theta <= -5.9, 1, 0), 
         #early_stop =ifelse(acc < 0.6, 1, 0), 
         true_low = as.numeric(thetaTrue ==-6)) %>% 
  ungroup() %>% 
  summarise(
    TP = sum(early_stop == 1 & true_low == 1),
    FP = sum(early_stop == 1 & true_low == 0),
    FN = sum(early_stop == 0 & true_low == 1),
    TN = sum(early_stop == 0 & true_low == 0)
  ) %>%
  mutate(
    precision = TP / (TP + FP),
    recall = TP / (TP + FN),
    f1 = 2 * precision * recall / (precision + recall),
    accuracy = (TP + TN) / (TP + FP + FN + TN)
  )
```

```{r}
hi.prob <- hi %>% 
  filter(trialNumTotal <= 30) %>% 
  group_by(pid, thetaTrue) %>% 
  summarise(acc = mean(resp)) %>% 
  mutate(prob_stop = ifelse(acc < 0.65, 1, 0)) %>% 
  group_by(thetaTrue) %>% 
  summarise(prob_stop_per_theta = mean(prob_stop))
  
ggplot(hi.prob, aes(x = thetaTrue, y = prob_stop_per_theta)) +
  geom_smooth(size = 0.5) +
  theme_minimal() +
  labs(x = "Theta Estimate", y = "prob")
```

```{r}
library(tidyverse)

# trial_vals <- c(10, 15, 20, 25, 30)
# theta_thresh <- c(-6, -5.5, -5)
# 
# params <- expand_grid(trialNumTotal = trial_vals,
#                       theta_threshold = theta_thresh)
# 
# hi.prob <- params %>%
#   pmap_dfr(function(trialNumTotal, theta_thresh) {
# 
#     hi %>%
#       filter(trialNumTotal <= !!trialNumTotal) %>%
#       group_by(pid, thetaTrue) %>%
#       summarise(acc = mean(resp), max_theta = max(thetaEstimate), .groups = "drop") %>% 
#       mutate(prob_stop = ifelse(max_theta <= theta_thresh, 1, 0)) %>%
#       group_by(thetaTrue) %>%
#       summarise(prob_stop_per_theta = mean(prob_stop), .groups = "drop") %>%
#       mutate(trialNumTotal = trialNumTotal,
#              theta_thresh = theta_thresh)
#   })

trial_vals <- c(20, 30, 40)
theta_thresh <- c(-6, -5.5, -5)

params <- expand_grid(trial = trial_vals,
                      theta_threshold = theta_thresh)

hi.prob <- params %>%
  pmap_dfr(function(trial, theta_threshold) {   # fix spelling here

    hi %>%
      filter(trialNumTotal <= trial, trialNumTotal > trial - 10) %>%       # just use argument directly
      group_by(pid, thetaTrue) %>%
      summarise(acc = mean(resp),
                max_theta = max(thetaEstimate),
                .groups = "drop") %>% 
      mutate(prob_stop = ifelse(max_theta <= theta_threshold, 1, 0)) %>%
      group_by(thetaTrue) %>%
      summarise(prob_stop_per_theta = mean(prob_stop), .groups = "drop") %>%
      mutate(trialNumTotal = trial,
             theta_threshold = theta_threshold)       # fix spelling
  })
hi.prob
ggplot(hi.prob,
       aes(x = thetaTrue,
           y = prob_stop_per_theta,
           color = factor(trialNumTotal),
           linetype = factor(theta_threshold))) +
  geom_smooth(se = FALSE, size = 0.8) +
  theme_minimal() +
  labs(x = "Theta Estimate",
       y = "Probability of Stop",
       color = "Trial Num",
       linetype = "Accuracy Threshold")

library(RColorBrewer)

cols <- brewer.pal(5, "Set1")

ggplot(hi.prob,
       aes(x = thetaTrue,
           y = prob_stop_per_theta,
           color = factor(trialNumTotal),
           )) +
  scale_color_manual(values = cols) +
  geom_smooth(se = FALSE, size = 0.8) +
  theme_minimal() +
  labs(x = "Theta Estimate",
       y = "Probability of Stop",
       color = "Trial Num",
       linetype = "Accuracy Threshold") + facet_wrap(~theta_threshold) + theme(legend.position = "bottom") + xlim(-6, 6)
```
```{r}
library(tidyverse)

trial_vals <- c(10, 15, 20, 25, 30)
acc_thresh <- c(0.5, 0.55, 0.6, 0.65, 0.7)

params <- expand_grid(trialNumTotal = trial_vals,
                      acc_threshold = acc_thresh)

hi.prob <- params %>%
  pmap_dfr(function(trialNumTotal, acc_threshold) {

    hi %>%
      filter(trialNumTotal <= !!trialNumTotal) %>%
      group_by(pid, thetaTrue) %>%
      summarise(acc = mean(resp), .groups = "drop") %>%
      mutate(prob_stop = ifelse(acc < acc_threshold, 1, 0)) %>%
      group_by(thetaTrue) %>%
      summarise(prob_stop_per_theta = mean(prob_stop), .groups = "drop") %>%
      mutate(trialNumTotal = trialNumTotal,
             acc_threshold = acc_threshold)
  })

ggplot(hi.prob,
       aes(x = thetaTrue,
           y = prob_stop_per_theta,
           color = factor(trialNumTotal),
           linetype = factor(acc_threshold))) +
  geom_smooth(se = FALSE, size = 0.8) +
  theme_minimal() +
  labs(x = "Theta Estimate",
       y = "Probability of Stop",
       color = "Trial Num",
       linetype = "Accuracy Threshold")

library(RColorBrewer)

cols <- brewer.pal(5, "Set1")

ggplot(hi.prob,
       aes(x = thetaTrue,
           y = prob_stop_per_theta,
           color = factor(trialNumTotal),
           )) +
  scale_color_manual(values = cols) +
  geom_smooth(se = FALSE, size = 0.8) +
  theme_minimal() +
  labs(x = "Theta Estimate",
       y = "Probability of Stop",
       color = "Trial Num",
       linetype = "Accuracy Threshold") + facet_wrap(~acc_threshold) + theme(legend.position = "bottom")
```

## Empirical Evidence: from hard item bank
```{r}
df.student.data <- read_csv("data/swr_with_94_trials.csv")
```


```{r}
df.student.data.k.2.clean <- df.student.data %>% 
  filter(completed == "TRUE", assessment_stage == "test_response") %>%
  filter(user_grade %in% c("1", "2", "Kindergarten")) 
  
```

```{r}
df.thetas <- df.student.data.k.2.clean %>% 
  filter(trial_num_total == 94) %>% 
  select(user_grade, run_id, theta_estimate, theta_std_err) %>% 
  rename(trueTheta = theta_estimate, trueSE = theta_std_err)
```

```{r}
hist(df.thetas$trueTheta)
```

### what patterns do the low performance share? 

```{r}
df.hi <- df.student.data.k.2.clean %>% 
  filter(trial_num_total < 30) %>% 
  #mutate(test_stage = (trial_num_total - 1) %/% 10 + 1 ) %>% 
  group_by(run_id) %>% 
  summarise(median_rt = median(rt)) %>% 
  left_join(df.thetas) %>% 
  filter(trueTheta < -5)

ggplot(df.hi, aes(x = median_rt)) +
  geom_density() +
  ylim(0, NA) + xlim(0, 5000)
```
```{r}
df.std %>% 
  group_by(run_id) %>% 
  tally()
```


```{r}
df.std <- df.student.data.k.2.clean %>% 
  left_join(df.thetas, by = "run_id") %>% 
  filter(trueTheta == -6, grade == "Kindergarten")

df.std %>% filter(run_id == "0BSpYthIwrL2xhanhqAg")

ggplot(df.std, aes(x = trial_num_total, theta_std_err, group = run_id)) +
  geom_line(alpha = 0.2) + xlim(12, 30) + ylim(0, 30)

ggplot(df.std %>% filter(run_id == "0BSpYthIwrL2xhanhqAg"), aes(x = trial_num_total, theta_std_err, group = run_id)) +
  geom_line(alpha = 0.2) + xlim(12, 80) + ylim(0, 30)

ggplot(df.std , aes(x = trial_num_total, theta_estimate, group = run_id)) +
  geom_line(alpha = 0.2) + xlim(12, 80) 

```

```{r}
df.30 <- df.student.data.k.2.clean %>% 
  filter(trial_num_total == 30) %>% 
  select(run_id, theta_estimate, theta_std_err)

df.e <- df.student.data.k.2.clean %>% 
  filter(trial_num_total <= 30) %>% 
  group_by(run_id) %>% 
  summarise(
    acc_20 = mean(correct, na.rm = TRUE),
    rt_20  = median(rt, na.rm = TRUE)
  ) %>% 
  left_join(df.thetas, by = "run_id") %>% 
  mutate(prob_stop = ifelse(acc_20 < 0.65, 1, 0)) %>% 
  left_join(df.30) %>% 
  filter(rt_20 < 2000)

ggplot(df.e ,
       aes(x = trueTheta,
           y = prob_stop, color = theta_std_err)) + geom_smooth() + xlim(-6, 6) + geom_point(alpha = 0.05)

ggplot(df.e ,
       aes(x = trueSE,
           y = theta_std_err, color = rt_20)) + geom_smooth()  + geom_point(alpha = 0.3) + ylim(0, 6) + coord_equal() + scale_color_viridis_c()
```
```{r}
ggplot(df.30,
       aes(x = theta_estimate,
           y = theta_std_err)) +
  geom_point() + ylim(0, 5)
```

TODO: get confusion matrix

```{r}
num_trials <- seq(20, 40, by = 5 )


df.e <- df.student.data.k.2.clean %>% 
  filter(trial_num_total <= num_trials, trial_num_total > 10) %>% 
  group_by(run_id) %>% 
  summarise(acc_20 = mean(correct), .groups = "drop") %>% 
  left_join(df.thetas, by = "run_id")

acc_thresh <- seq(0.5, 0.70, by = 0.05)

num_trials <- seq(20, 40, by = 5)

conditions <- expand_grid(
  acc_thresh = acc_thresh,
  num_trials = num_trials
)

results <- df.e %>%
  crossing(conditions) %>%
  mutate(
    prob_stop = as.numeric(acc_20 < acc_thresh),
    true_low = as.numeric(theta_estimate < -5)
  ) %>%
  group_by(acc_thresh,num_trials) %>%
  summarise(
    TP = sum(prob_stop & true_low),
    FP = sum(prob_stop & !true_low),
    FN = sum(!prob_stop & true_low),
    TN = sum(!prob_stop & !true_low),

    precision = TP / (TP + FP),
    recall = TP / (TP + FN),
    f1 = 2 * precision * recall / (precision + recall),
    accuracy = (TP + TN) / (TP + FP + FN + TN),

    # AUC computed for this threshold combination
    auc = as.numeric(auc(roc(true_low, prob_stop, quiet = TRUE))),

    .groups = "drop"
  )

ggplot(results,
       aes(x = theta_thresh,
           y = acc_thresh,
           fill = f1)) +
  geom_tile() +
  scale_fill_viridis_c(name = "F1 score") +
  theme_minimal() +
  labs(
    x = "Theta threshold",
    y = "Accuracy threshold"
  ) 
```

```{r}
library(tidyverse)
library(pROC)

#---------------------------------------
# 1. Condition grid
#---------------------------------------

theta_thresh <- seq(-6, -5, by = 0.02)
num_trials <- seq(30, 50, by = 2)

conditions <- expand_grid(
  theta_thresh = theta_thresh,
  num_trials = num_trials
)

#---------------------------------------
# 2. Precompute cumulative accuracy
#---------------------------------------

df.acc <- df.student.data.k.2.clean %>%
  filter(trial_num_total > 10) 

#---------------------------------------
# 3. Compute accuracy up to each trial cutoff
#---------------------------------------

# df.e <- df.acc %>%
#   crossing(num_trials) %>%
#   filter(trial_num_total <= num_trials) %>%
#   group_by(run_id, num_trials) %>%
#   summarise(acc_k = mean(correct), .groups = "drop") %>%
#   left_join(df.thetas, by = "run_id") %>% 
#   left_join(df.acc)

df.e <- df.acc %>%
  crossing(num_trials) %>%
  filter(trial_num_total == num_trials) %>%
  left_join(df.thetas, by = "run_id")

#---------------------------------------
# 4. Evaluate stopping rules
#---------------------------------------

# results <- df.e %>%
#   crossing(theta_thresh) %>%
#   mutate(
#     prob_stop = as.numeric(theta_estimate <= theta_thresh),
#     true_low = as.numeric(trueTheta < -5.5)
#   ) %>%
#   group_by(num_trials, theta_thresh) %>%
#   summarise(
#     TP = sum(prob_stop & true_low),
#     FP = sum(prob_stop & !true_low),
#     FN = sum(!prob_stop & true_low),
#     TN = sum(!prob_stop & !true_low),
# 
#     precision = TP / (TP + FP),
#     recall = TP / (TP + FN),
#     f1 = 2 * precision * recall / (precision + recall),
#     accuracy = (TP + TN) / (TP + FP + FN + TN),
# 
# 
#     .groups = "drop"
#   )

library(dplyr)
library(slider)

results <- df.e %>%
  crossing(theta_thresh) %>%
  arrange(run_id, trial_num_total) %>%  # ensure correct order
  group_by(run_id, num_trials, theta_thresh) %>%
  mutate(
    # rolling check: are ALL last 5 theta_estimates <= threshold?
    last5_below = slide_lgl(
      theta_estimate,
      ~ all(.x <= theta_thresh),
      .before = 4,   # include current + 4 previous = 5 trials
      .complete = TRUE
    ),
    
    prob_stop = as.numeric(last5_below),
    true_low = as.numeric(trueTheta < -5.5)
  ) %>%
  filter(trial_num_total == num_trials) %>%  # evaluate at stopping point
  group_by(num_trials, theta_thresh) %>%
  summarise(
    TP = sum(prob_stop & true_low),
    FP = sum(prob_stop & !true_low),
    FN = sum(!prob_stop & true_low),
    TN = sum(!prob_stop & !true_low),

    precision = TP / (TP + FP),
    recall = TP / (TP + FN),
    f1 = 2 * precision * recall / (precision + recall),
    accuracy = (TP + TN) / (TP + FP + FN + TN),

    .groups = "drop"
  )

ggplot(results,
       aes(x = num_trials,
           y = theta_thresh,
           fill = precision)) +
  geom_tile(color = "white") +
  scale_fill_viridis_c(name = "precision") +
  theme_minimal(base_size = 13) +
  labs(
    x = "Number of trials used",
    y = "Accuracy threshold"
  )

ggplot(results,
       aes(x = num_trials,
           y = theta_thresh,
           fill = accuracy)) +
  geom_tile(color = "white") +
  scale_fill_viridis_c(name = "accuracy") +
  theme_minimal(base_size = 13) +
  labs(
    x = "Number of trials used",
    y = "Accuracy threshold"
  )
```
```{r}
df.cm <- df.e %>%
  filter(trial_num_total > 30, trial_num_total <= 50) %>% 
  group_by(run_id, trueTheta) %>% 
  summarise(max_theta = max(theta_estimate),  .groups = "drop") %>% 
  mutate(
    early_stop = ifelse(max_theta <= -5.9, 1, 0),
    true_low = as.numeric(trueTheta <= -6)
  )

df.cm %>%
  summarise(
    TP = sum(early_stop == 1 & true_low == 1),
    FP = sum(early_stop == 1 & true_low == 0),
    FN = sum(early_stop == 0 & true_low == 1),
    TN = sum(early_stop == 0 & true_low == 0)
  ) %>%
  mutate(
    precision = TP / (TP + FP),
    recall = TP / (TP + FN),
    f1 = 2 * precision * recall / (precision + recall),
    accuracy = (TP + TN) / (TP + FP + FN + TN)
  )
  
```

```{r}

df.e <- df.acc %>%
  group_by(run_id) %>%
  # Check if the current AND previous 4 trials are all below the threshold
  # We use slide_lgl to see if the condition 'theta_estimate <= theta_thresh' 
  # holds true for the entire window of 5.
  crossing(theta_thresh) %>%
  mutate(
    # This creates a logical vector: TRUE if the last 5 trials met the criteria
    met_stop_condition = slide_lgl(
      theta_estimate, 
      ~all(.x <= theta_thresh), 
      .before = 4, 
      .complete = TRUE
    )
  ) %>%
  ungroup() %>%
  # Now filter to the specific num_trials you want to evaluate
  crossing(num_trials) %>%
  filter(trial_num_total == num_trials) %>%
  left_join(df.thetas, by = "run_id")

#---------------------------------------
# 4. Evaluate stopping rules
#---------------------------------------

results <- df.e %>%
  mutate(
    # Use the rolling condition instead of the single point check
    prob_stop = as.numeric(met_stop_condition),
    true_low  = as.numeric(trueTheta <=-6)
  ) %>%
  # Remove NAs (the first 4 trials of any run cannot have a "last 5" window)
  filter(!is.na(prob_stop)) %>%
  group_by(num_trials, theta_thresh) %>%
  summarise(
    TP = sum(prob_stop & true_low, na.rm = TRUE),
    FP = sum(prob_stop & !true_low, na.rm = TRUE),
    FN = sum(!prob_stop & true_low, na.rm = TRUE),
    TN = sum(!prob_stop & !true_low, na.rm = TRUE),

    precision = TP / (TP + FP),
    recall    = TP / (TP + FN),
    f1        = 2 * precision * recall / (precision + recall),
    accuracy  = (TP + TN) / (TP + FP + FN + TN),

    .groups = "drop"
  )
```
```{r}
results 
```

```{r}
library(tidyverse)
library(pROC)

#---------------------------------------
# 1. Condition grid
#---------------------------------------

acc_thresh <- seq(-6, -3, by = 0.2)
num_trials <- seq(30, 40, by = 2)

conditions <- expand_grid(
  acc_thresh = acc_thresh,
  num_trials = num_trials
)

#---------------------------------------
# 2. Precompute cumulative accuracy
#---------------------------------------

df.acc <- df.student.data.k.2.clean %>%
  filter(trial_num_total > 10) 

#---------------------------------------
# 3. Compute accuracy up to each trial cutoff
#---------------------------------------

df.e <- df.acc %>%
  crossing(num_trials) %>%
  filter(trial_num_total == num_trials) %>%
  left_join(df.thetas, by = "run_id")

```
```{r}
#---------------------------------------
# 4. Evaluate stopping rules
#---------------------------------------

results <- df.e %>%
  crossing(acc_thresh) %>%
  mutate(
    prob_stop = as.numeric(theta_estimate < acc_thresh),
    true_low = as.numeric(trueTheta < -5)
  ) %>%
  group_by(num_trials, acc_thresh) %>%
  summarise(
    TP = sum(prob_stop & true_low),
    FP = sum(prob_stop & !true_low),
    FN = sum(!prob_stop & true_low),
    TN = sum(!prob_stop & !true_low),

    precision = TP / (TP + FP),
    recall = TP / (TP + FN),
    f1 = 2 * precision * recall / (precision + recall),
    accuracy = (TP + TN) / (TP + FP + FN + TN),
    auc = as.numeric(auc(roc(true_low, theta_estimate, quiet = TRUE))),
    .groups = "drop"
  )

ggplot(results,
       aes(x = num_trials,
           y = acc_thresh,
           fill = precision)) +
  geom_tile(color = "white") +
  scale_fill_viridis_c(name = "f1") +
  theme_minimal(base_size = 13) +
  labs(
    x = "Number of trials used",
    y = "Accuracy threshold"
  )

ggplot(results,
       aes(x = num_trials,
           y = acc_thresh,
           fill = auc)) +
  geom_tile(color = "white") +
  scale_fill_viridis_c(name = "f1") +
  theme_minimal(base_size = 13) +
  labs(
    x = "Number of trials used",
    y = "Accuracy threshold"
  )
```

```{r}
results
```

